---
title: 'STAT 1341 Midterm Project: COVID-19 and the NBA'
author: "Chris Park and Kevin Zhang"
output:
  html_document:
    df_print: paged
---

# Data Cleaning

```{r}
library(readxl)

# 2019-2020 NBA Player Stats
rs_player_19_20 = read_excel('2019-2020 NBA Player Stats Regular Season.xlsx')

# 2020-2021 NBA Player Stats
rs_player_20_21 = read_excel('2020-2021 NBA Player Stats Regular Season.xlsx')

# 2018-2019 NBA Team Stats
team_18_19 = read_excel('2018-2019 NBA Team Stats.xls')

# 2019-2020 NBA Team Stats
team_19_20 = read_excel('2019-2020 NBA Team Stats.xls')

# 2020-2021 NBA Team Stats
team_20_21 = read_excel('2020-2021 NBA Team Stats.xls')

# Dropping unwanted columns in player stats datasets
rs_player_19_20 = subset(rs_player_19_20, select = -c(RANK,POS,AGE,MPG,MIN_percentage,FTA,two_PA,twoP_percentage,three_PA,threeP_percentage,TRB_percentage,AST_percentage))

rs_player_20_21 = subset(rs_player_20_21, select = -c(RANK,POS,AGE,MPG,MIN_percentage,FTA,two_PA,twoP_percentage,three_PA,threeP_percentage,TRB_percentage,AST_percentage))

# Dropping unwanted columns in team stats datasets 
team_18_19 = subset(team_18_19, select = -c(Rk,Age,PW,PL,MOV,SOS,SRS,NRtg,FTr,three_PAr,18,23,eFG_per,TOV_per,DRB_per,FT_FGA_d,28,Arena,Attend.))
team_19_20 = subset(team_19_20, select = -c(Rk,Age,PW,PL,MOV,SOS,SRS,NRtg,FTr,three_PAr,18,23,eFG_per,TOV_per,DRB_per,FT_FGA_d,28,Arena,Attend.))
team_20_21 = subset(team_20_21, select = -c(Rk,Age,PW,PL,MOV,SOS,SRS,NRtg,FTr,three_PAr,18,23,eFG_per,TOV_per,DRB_per,FT_FGA_d,28,Arena,Attend.))

# Dropping league average row in team stats datasets
team_18_19 = team_18_19[-c(31),]
team_19_20 = team_19_20[-c(31),]
team_20_21 = team_20_21[-c(31),]

# Changing row with NA to 0 for attendance
team_20_21[is.na(team_20_21)] = 0

# Dropping rows that have NA
rs_player_19_20 = na.omit(rs_player_19_20)
rs_player_20_21 = na.omit(rs_player_20_21)

# Removing duplicates: if the player was on two teams, the team that the player had more games played is kept
rs_player_19_20 = rs_player_19_20[!duplicated(rs_player_19_20$`FULL NAME`),]
rs_player_20_21 = rs_player_20_21[!duplicated(rs_player_20_21$`FULL NAME`),]

# Removing players that did not play in both seasons
rs_player_19_20 = subset(rs_player_19_20, rs_player_19_20$`FULL NAME` %in% rs_player_20_21$`FULL NAME`)
rs_player_20_21 = subset(rs_player_20_21, rs_player_20_21$`FULL NAME` %in% rs_player_19_20$`FULL NAME`)
```

# Descriptive Statistics

## Comparing Player Statistics

### Summary Statistics: consisting of N, Mean, Std. Dev., Min, Pctl.25, Pctl.75, and Max
```{r}
library(vtable)
st(rs_player_19_20)
st(rs_player_20_21)
```

### Plotting and Correlations: Checking for any noteworthy relationships
```{r}
# Turnover Percentage
plot(rs_player_19_20$TO_percentage, rs_player_20_21$TO_percentage, xlab="19-20 TO%", ylab="20-21 TO%")
cor(rs_player_19_20$TO_percentage, rs_player_20_21$TO_percentage)

# Free Throw Percentage
plot(rs_player_19_20$FT_percentage, rs_player_20_21$FT_percentage, xlab="19-20 FT%", ylab="20-21 FT%")
cor(rs_player_19_20$FT_percentage, rs_player_20_21$FT_percentage)

# Effective Field Goal Percentage
plot(rs_player_19_20$eFG_percentage, rs_player_20_21$eFG_percentage, xlab="19-20 eFG%", ylab="20-21 eFG%")
cor(rs_player_19_20$eFG_percentage, rs_player_20_21$eFG_percentage)

# True Shooting Percentage
plot(rs_player_19_20$TS_percentage, rs_player_20_21$TS_percentage, xlab="19-20 TS%", ylab="20-21 TS%")
cor(rs_player_19_20$TS_percentage, rs_player_20_21$TS_percentage)

# Points Per Game
plot(rs_player_19_20$PPG, rs_player_20_21$PPG, xlab="19-20 PPG", ylab="20-21 PPG")
cor(rs_player_19_20$PPG, rs_player_20_21$PPG)

# Rebounds Per Game
plot(rs_player_19_20$RPG, rs_player_20_21$RPG, xlab="19-20 RPG", ylab="20-21 RPG")
cor(rs_player_19_20$RPG, rs_player_20_21$RPG)

# Assists Per Game
plot(rs_player_19_20$APG, rs_player_20_21$APG, xlab="19-20 APG", ylab="20-21 APG")
cor(rs_player_19_20$APG, rs_player_20_21$APG)

# Steals Per Game
plot(rs_player_19_20$SPG, rs_player_20_21$SPG, xlab="19-20 SPG", ylab="20-21 SPG")
cor(rs_player_19_20$SPG, rs_player_20_21$SPG)

# Blocks Per Game
plot(rs_player_19_20$BPG, rs_player_20_21$BPG, xlab="19-20 BPG", ylab="20-21 BPG")
cor(rs_player_19_20$BPG, rs_player_20_21$BPG)

# Turnovers Per Game
plot(rs_player_19_20$TOPG, rs_player_20_21$TOPG, xlab="19-20 TOPG", ylab="20-21 TOPG")
cor(rs_player_19_20$TOPG, rs_player_20_21$TOPG)

# ViVersatility Index
plot(rs_player_19_20$VIVersatility, rs_player_20_21$VIVersatility, xlab="19-20 VIV", ylab="20-21 VIV")
cor(rs_player_19_20$VIVersatility, rs_player_20_21$VIVersatility)

# Offensive Rating
plot(rs_player_19_20$ORTG, rs_player_20_21$ORTG, xlab="19-20 ORTG", ylab="20-21 ORTG")
cor(rs_player_19_20$ORTG, rs_player_20_21$ORTG)

# Defensive Rating
plot(rs_player_19_20$DRTG, rs_player_20_21$DRTG, xlab="19-20 DRTG", ylab="20-21 DRTG")
cor(rs_player_19_20$DRTG, rs_player_20_21$DRTG)
```

## Comparing Team Statistics

### Summary Statistics: consisting of N, Mean, Std. Dev., Min, Pctl.25, Pctl.75, and Max
```{r}
library(vtable)
st(team_19_20)
st(team_20_21)
```

### Plotting, Correlations, Histograms: Checking for any noteworthy relationships
```{r}
# Wins
plot(team_19_20$W, team_20_21$W, xlab="19-20 Wins", ylab="20-21 Wins")
cor(team_19_20$W, team_20_21$W)
hist(team_19_20$W)
hist(team_20_21$W)

# Losses
plot(team_19_20$L, team_20_21$L, xlab="19-20 Losses", ylab="20-21 Losses")
cor(team_19_20$L, team_20_21$L)
hist(team_19_20$L)
hist(team_20_21$L)

# Offensive Rating
plot(team_19_20$ORtg, team_20_21$ORtg, xlab="19-20 ORTG", ylab="20-21 ORTG")
cor(team_19_20$ORtg, team_20_21$ORtg)
hist(team_19_20$ORtg)
hist(team_20_21$ORtg)

# Defensive Rating
plot(team_19_20$DRtg, team_20_21$DRtg, xlab="19-20 DRTG", ylab="20-21 DRTG")
cor(team_19_20$DRtg, team_20_21$DRtg)
hist(team_19_20$DRtg)
hist(team_20_21$DRtg)

# Pace
plot(team_19_20$Pace, team_20_21$Pace, xlab="19-20 Pace", ylab="20-21 Pace")
cor(team_19_20$Pace, team_20_21$Pace)
hist(team_19_20$Pace)
hist(team_20_21$Pace)

# True Shooting Percentage
plot(team_19_20$TS_percentage, team_20_21$TS_percentage, xlab="19-20 TS%", ylab="20-21 TS%")
cor(team_19_20$TS_percentage, team_20_21$TS_percentage)
hist(team_19_20$TS_percentage)
hist(team_20_21$TS_percentage)

# Effective Field Goal Percentage
plot(team_19_20$eFG_percentage, team_20_21$eFG_percentage, xlab="19-20 eFG%", ylab="20-21 eFG%")
cor(team_19_20$eFG_percentage, team_20_21$eFG_percentage)
hist(team_19_20$eFG_percentage)
hist(team_20_21$eFG_percentage)

# Turnover Percentage
plot(team_19_20$TOV_percentage, team_20_21$TOV_percentage, xlab="19-20 TOV%", ylab="20-21 TOV%")
cor(team_19_20$TOV_percentage, team_20_21$TOV_percentage)
hist(team_19_20$TOV_percentage)
hist(team_20_21$TOV_percentage)

# Offensive Rebound Percentage
plot(team_19_20$ORB_percentage, team_20_21$ORB_percentage, xlab="19-20 ORB%", ylab="20-21 ORB%")
cor(team_19_20$ORB_percentage, team_20_21$ORB_percentage)
hist(team_19_20$ORB_percentage)
hist(team_20_21$ORB_percentage)

# Free Throws/Field Goal Attempts
plot(team_19_20$FT_FGA, team_20_21$FT_FGA, xlab="19-20 FT/FGA", ylab="20-21 FT/FGA")
cor(team_19_20$FT_FGA, team_20_21$FT_FGA)
hist(team_19_20$FT_FGA)
hist(team_20_21$FT_FGA)

# Attendance Per Game
plot(team_19_20$Attend._G, team_20_21$Attend._G, xlab="19-20 Attendance/G", ylab="20-21 Attendance/G")
cor(team_19_20$Attend._G, team_20_21$Attend._G)
hist(team_19_20$Attend._G)
hist(team_20_21$Attend._G)
```

# Inferential Statistics

## Multiple Linear Regression Models
```{r}
# 2018-2019 Team Stats Model
model1 = lm(W ~ ORtg + DRtg + Pace + TS_percentage + eFG_percentage + TOV_percentage + ORB_percentage + FT_FGA, data = team_18_19)
summary(model1)

# 2019-2020 Team Stats Model
model2 = lm(W ~ ORtg + DRtg + Pace + TS_percentage + eFG_percentage + TOV_percentage + ORB_percentage + FT_FGA, data = team_19_20)
summary(model2)

# 2020-2021 Team Stats Model
model3 = lm(W ~ ORtg + DRtg + Pace + TS_percentage + eFG_percentage + TOV_percentage + ORB_percentage + FT_FGA, data = team_20_21)
summary(model3)
```


